Seasonality in governing upper-ocean dynamics at submesoscales

### 2016-06-04 writeup/rocha_etal.tex
Updates figures
Adds new figure

\authors{Cesar, Sarah, Teri, \& Bill \altaffilmark{1}}
% Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
% John B. McDougall\altaffilmark{4}, and S. Visconti\altaffilmark{5}}

@@

\begin{abstract}
(Type abstract here)
\end{abstract}

@@

The horizontal divergent part of the flow is significant: the RMS is $\delta/f > 0.1$).
Most of these horizontal divergent flows are supra-inertial and project
on scales smaller than 100 km: daily-averaging or smoothing the velocity fields
 (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
Interestingly, the seasonal cycle in horizontal divergence is out-of-phase with the
seasonal cycle of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
Summer/early Fall than it is in late Winter/early Spring. The divergence associated
with submesoscale subinertial eddies is small and its seasonal cycle is in phase
with that of vorticity (see green lines).  The seasonal cycle of the horizontal
divergence is roughly in phase with the seasonal cycle of the near-surface (0-50m)
stratification.
\begin{figure*}[ht]
\begin{center}
\includegraphics[width=1.\textwidth]{figs/fig1.pdf}
 \caption{Snapshots of surface vorticity (a and b) and a transect
          of potential density at 165$^\circ$E (c and d). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
 \caption{Surface (horizontal) KE frequency spectra. Solid lines
 are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
 based on 100-km-smoothed fields.}
 \label{fig4}
 \end{center}
\end{figure}

@@

%  this should go in a supplementary material
\appendix
\section{How well does the model captures high-frequency modes?}
To assess how well the LLC2160 represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

@@

@@

\begin{acknowledgments}
Thanks to D. Menemenlis (JPL/NASA) and the other MITgcm usual suspects.
This research was funded by NSF and NASA.
\end{acknowledgments}

>>>>

\authors{Cesar et al. \altaffilmark{1}}
% Eric Brown,\altaffilmark{1,2} Rick Williams,\altaffilmark{3}
% John B. McDougall\altaffilmark{4}, and S. Visconti\altaffilmark{5}}

@@

\begin{abstract}
    We show that internal waves are a loud noise.
\end{abstract}

@@

The horizontal divergent part of the flow is significant: the RMS is $\delta/f > 0.1$).
Most of these horizontal divergent flows are supra-inertial and project
on scales smaller than 100 km: daily-averaging or smoothing the velocity fields
 (100 km cutoff) reduces the horizontal divergence by about 70$\%$.
Interestingly, the seasonal cycle in horizontal divergence is out-of-phase with the
seasonal cycle of vertical vorticity: the horizontal divergence is about 30$\%$ larger in late
Summer/early Fall than it is in late Winter/early Spring. The divergence associated
with submesoscale subinertial eddies is small and its seasonal cycle is in phase
with that of vorticity (see green lines).  The seasonal cycle of the horizontal
divergence is roughly in phase with the seasonal cycle of the near-surface (0-50m)
stratification.
\begin{figure*}[ht]
\begin{center}
\hspace{-1.25cm}\includegraphics[width=.7\textwidth]{figs/fig1_1.pdf}\\
\vspace{-.125cm}
\includegraphics[width=.9\textwidth]{figs/fig1_2.pdf}
 \caption{(a) The study region with the subregion where the LLC outputs are
          analyzed. Snapshots of surface vorticity (b and c) and a transect
          of potential density at 165$^\circ$E (d and e). The snapshots were
          taken at 00:00 UTC.}
\vspace{-1.5cm}
 \label{fig1}
 \end{center}
 \end{figure*}

@@

%\begin{figure}[ht]
%  \begin{center}
%    \includegraphics[width=.45\textwidth]{figs/fig4.pdf}
% \caption{Surface (horizontal) KE frequency spectra. Solid lines
% are spectra based on the 1/24$^\circ$ fields, dashed line are spectra
% based on 100-km-smoothed fields.}
% \label{fig4}
% \end{center}
%\end{figure}

@@

%  this should go in a supplementary material
\clearpage

\section{Supplemental material 1: Spectral errors}

\section{Supplemental material 2: On the LLC simulations}

\subsection{Seasonality in the upper-ocean stratification}

\subsection{How well does the model captures high-frequency modes?}
To assess how well the LLC2160 represents the high-frequency variability,
we compare the model near-surface velocity field against two available
mooring data: KEO (the data is available online at \texttt{http://www.pmel.noaa.gov/ocs/KEO})
and KESS mooring $\#$7 (the data is available online at \texttt{http://uskess.whoi.edu}).
For proper comparison, we use model data closest to these two moorings (west of
the region considered in the present study). Focusing on high-frequencies, we
calculate frequency spectral estimates every month. There are about 48 independent
spectral realizations.

@@

\section{Supplemental material 3: Seasonality in pressure modes near-surface amplitude}

@@

\begin{acknowledgments}
Thanks to the MITgcm usual suspects.
This research was funded by NSF and NASA.
\end{acknowledgments}
